Bayesian statistics and swine flu

### 2009-09-25 docs/pandemic/bayesPandemic.rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@412 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\comment{
<<setup,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=54)
# source in all the R files
library(pandemic)
@
}

@@

params = pandemicParams( 
	InfOns = c(mean=1.5, shape=2, zeros=.1),
	OnsMedM= c(mean=1, shape=1, zeros=.1 ),
	OnsMedS= c(mean=1, shape=1, zeros=.1 ),
	OnsMedD= c(mean=1, shape=1, zeros=.1 ),
	MedRec = c(mean=3, shape=2, zeros=.1, lost=0.95 ),
  MedHospS = c(mean = 2, shape = .5, zeros = 0.1),
  MedHospD = c(mean = 2, shape = 2, zeros = 0.1),
	HospRec = c(mean=3, shape=1.5, zeros=.01 ),
	HospDeath = c(mean=5, shape=0.8, zeros=.05 ),
ageProbs=list(
	S=addAgeProbs(ageSeq, 0.02+0.8*dnorm(ageSeq, mean=20, sd=10)),
	D=addAgeProbs(ageSeq, 0.01+0.5*dnorm(ageSeq, mean=30, sd=5))  ),
	probs=NULL
	)
tempList = (params[c("InfOns","OnsMedM","OnsMedS", "OnsMedD", "MedRec","MedHospS","MedHospD","HospRec","HospDeath")])
temp = unlist(tempList)
temp = temp [-grep("lost",names(temp))]
tempMat = matrix(temp, ncol=4, byrow=T)
rownames(tempMat) = names(tempList)
colnames(tempMat) = names(tempList[[1]])
paramMat = tempMat[,c("mean","shape","zeros")]
library(Hmisc)
@

@@

data = simEpidemic(params, delta=20, days=28)

table(data$observedType)
@
<<fig=false>>=
toshow = NULL
for(Dshow in levels(data$observedType))
	toshow = c(toshow, which(data$observedType==Dshow)[1:2])
toshow = na.omit(toshow)	
@
<<simData2,echo=true,fig=false>>=
data[toshow,-c(1,7)]
@

@@

@@

\begin{frame}[fragile]
\frametitle{Results}
<<mcmcResults,echo=true,fig=false>>=
scale = mcmcScale(params, 0.075,minScale = 0.05 )
paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=60, thin = 5)

\begin{frame}
\frametitle{Fatality probability}
\begin{columns}
\column{0.5\textwidth}
\begin{block}{trace plots}
<<tracePlot,fig=true>>=
plot(paramSample[,"probs.D"], type="l")
abline(h=params$probs["D"], col="red")
@
\end{block}
\column{0.5\textwidth}

\begin{block}{Histogram}
<<histPost>>=
hist(paramSample[,"probs.D"], breaks=12)
abline(v=params$probs["D"], col="red")
@
\end{block}
\end{columns}
\end{frame}

postSummary[1:12,-c(1,4)]
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Hospital demand}
<<hospLoad,echo=T>>=
# generate hospital times
hospSample = simHospitals(paramSample , Ndays=20, NinfectionsPerDay=20)

>>>>

\comment{
<<setup,fig=false>>=
options(SweaveHooks=list(fig=function() par(mar=c(2.5,2.5,0.1,0.1), mgp=c(1.5, 0.5, 0), cex=0.6)), width=54, digits=3)
# source in all the R files
library(pandemic)
@
}

@@

params = pandemicParams( 
	InfOns = c(mean=1.5, shape=2, zeros=.1),
	OnsMedM= c(mean=1, shape=1, zeros=.1 ),
	OnsMedS= c(mean=1, shape=1, zeros=.1 ),
	OnsMedD= c(mean=1, shape=1, zeros=.1 ),
	MedRec = c(mean=3, shape=2, zeros=.1, lost=0.95 ),
  MedHospS = c(mean = 2, shape = .5, zeros = 0.1),
  MedHospD = c(mean = 2, shape = 2, zeros = 0.1),
	HospRec = c(mean=3, shape=1.5, zeros=.01 ),
	HospDeath = c(mean=5, shape=0.3, zeros=.05 ),
ageProbs=list(
	S=addAgeProbs(ageSeq, 0.02+1.5*dnorm(ageSeq, mean=20, sd=10)),
	D=addAgeProbs(ageSeq, 0.01+dnorm(ageSeq, mean=30, sd=8))  ),
	probs=NULL
	)
tempList = (params[c("InfOns","OnsMedM","OnsMedS", "OnsMedD", "MedRec","MedHospS","MedHospD","HospRec","HospDeath")])
temp = unlist(tempList)
temp = temp [-grep("lost",names(temp))]
tempMat = matrix(temp, ncol=4, byrow=T)
rownames(tempMat) = names(tempList)
colnames(tempMat) = names(tempList[[1]])
paramMat = tempMat[,c("mean","shape","zeros")]
library(Hmisc)
@

@@

data = simEpidemic(params, delta=50, days=42)

table(data$observedType)
@
<<fig=false>>=
toshow = NULL
for(Dshow in levels(data$observedType))
	toshow = c(toshow, which(data$observedType==Dshow)[1:2])
toshow = na.omit(toshow)	
#$
@
<<simData2,echo=true,fig=false>>=
data[toshow,-c(1,7)]
@

@@

\begin{frame}[fragile]
\frametitle{Fitting the model}
(to the simulated data)
<<mcmcResults,echo=true,fig=false>>=
scale = mcmcScale(params, 0.075,minScale = 0.05 )
paramSample = mcmcPandemic(data,params,priors,sigma=scale,runs=100, thin = 5)

\begin{frame}
\frametitle{Fatality probability, age 20}
\begin{columns}
\column{0.5\textwidth}
\begin{block}{trace plots}
<<tracePlot,fig=true>>=
toplot = "ageProbs.D.prob20"
plot(paramSample[,toplot], type="l")
abline(h=params$ageProbs$D[20,"prob"], col="red")
@
\end{block}
\column{0.5\textwidth}

\begin{block}{Histogram}
<<histPost>>=
hist(paramSample[,toplot], breaks=12)
abline(v=params$ageProbs$D[20,"prob"], col="red")
@
\end{block}
\end{columns}
\end{frame}

postSummary[grep("^(OnsMed|MedHosp)(M|S|D).(mean|shape)$",#$

rownames(postSummary)),-c(1,4)]
@
\end{frame}

\begin{frame}
\frametitle{Fatality rate}
<<fatalityAgeGraph>>=
ageCols = grep("^ageProbs.D.prob.[[:digit:]]+$", colnames(paramSample), value=T)
Sage= as.numeric(gsub("^ageProbs.D.prob", "", ageCols))

matplot(Sage, t(paramSample[,ageCols]), type="l", col=2:8, ylim=quantile(paramSample[,ageCols], probs=c(0.01, 0.99))) 
lines(Sage, apply(paramSample[,ageCols], 2, mean), lwd=3, col="black")
#$
@

\begin{frame}
\frametitle{Hospitalization rate}
<<fatalityAgeGraph>>=
ageColsS = grep("^ageProbs.S.prob.[[:digit:]]+$", colnames(paramSample), value=T)

theRate = paramSample[,ageCols] + (1-paramSample[,ageCols])*paramSample[,ageColsS]

matplot(Sage, t(theRate), type="l", col=2:8, ylim=quantile(theRate, probs=c(0.01, 0.99))) 
lines(Sage, apply(theRate, 2, mean), lwd=3, col="black")
#$
@

@@

\begin{frame}[fragile]
\frametitle{Hospital demand}
<<hospLoad,echo=T>>=
# generate hospital times
hospSample = simHospitals(paramSample , Ndays=60, 
	NinfectionsPerDay=c(seq(10, 100, len=30),seq(100, 10, len=30) ) )
